---
title: "Fitting parametric survival models in R"
author: "Roche"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: references.bib
csl: biomedicine.csl
vignette: >
  %\VignetteIndexEntry{Fitting parametric survival models in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 6,
  message = FALSE
)
```

<style type="text/css">
  
  body{ /* Normal  */
      font-size: 14px;
  }
td {  /* Table  */
    font-size: 10px;
}
h1.title {
  font-size: 38px;
}
h1 { /* Header 1 */
    font-size: 28px;
}
h2 { /* Header 2 */
    font-size: 22px;
}
h3 { /* Header 3 */
    font-size: 18px;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>
  
# Introduction
Parametric survival models are often the preferred method of extrapolating
survival data for use in economic models. The National Institute for Health and
Care Excellence (NICE) Decision Support Unit (DSU) technical support document
(TSD) 14 recommends that the Exponential, Weibull, Gompertz, log-logistic, log
normal and Generalized Gamma parametric models should all be
considered.[@latimer2011nice] More recently, NICE also discusses more flexible
models in NICE DSU TSD 21, however, more these models are not in the scope of
this package.[@rutherford2020nice] The Canadian Agency for Drugs and
Technologies in Health (CADTH) additionally specifies that the Gamma
distribution must also be considered. This document therefore details the
characteristics of each of these distributions and demonstrates how the
parameters from each distribution, outputted using the flexsurvPlus package, can
be implemented within an economic model.[@cadth2020] The Generalized F
distribution is not commonly used, however it has been included in this package
in case it is required.

The flexsurvPlus package allows the inclusion of a treatment effect in the
following four ways:
  
  * Separate models - Models fitted to each treatment arm separately

*  Independent shape models - Models fitted to both/all treatment arms including
a treatment covariate to model the effect of treatment on both the scale and
shape parameter(s) of the distribution

*  Common shape models - Models fitted to both/all treatment arms including a
treatment covariate to model the effect of treatment on the scale parameter of
the distribution. The shape parameter(s) of the distribution is common across
treatments which reflects an assumption of proportional hazards or accelerated
failure time between treatments depending on the distribution

*  One arm models - Models fitted to the entire dataset (no treatment strata)

This document details how to use the <tt>flexsurvPlus</tt> package to perform
these models. A separate vignette; "Parametric survival analysis using the
flexsurvPlus package: understanding the theory" details the theory behind the
models.


# Set up packages and data

## Install packages

The following packages are required to run this example:
  
```{r setup}
# Libraries
library(flexsurvPlus)
library(flexsurv)
library(tibble)
library(dplyr)
library(survival)
library(tidyr)
library(boot)

# Non-scientific notation
options(scipen=999) 

# Colours for plots
blue = rgb(69, 185, 209, max=255)
pink = rgb(211,78,147,max=255)
Dyellow = rgb(214, 200, 16, max=255)
orange<-rgb(247,139,21,max=255)

```

## Read in the data

To perform survival analyses, patient level data is required for the survival
endpoints.

This example uses a standard simulated data set (adtte). There is no standard
naming that is needed for this package however, there are some set variables
that are needed:
  
* Time - a numeric variable
* Event - a binary variable (event=1, censor=0)
* Treatment - a character variable with the name of the intervention treatment 

The data must be in "wide" format such that there is one row per patient and
columns for each endpoint separately. In this example, we analyze
progression-free survival (PFS).
```{r}
adtte <- sim_adtte(seed = 2020, rho = 0.6)
head(adtte)

# subset PFS data and rename
PFS_data <- adtte %>%
  filter(PARAMCD=="PFS") %>%
  transmute(USUBJID,
            ARMCD,
            PFS_days = AVAL,
            PFS_event = 1- CNSR
  )
head(PFS_data)

```

# Exploratory analysis
Before performing any statistical analysis, it is important to explore the data.
Most importantly is a Kaplan-Meier plot. A simple Kaplan-Meier plot has been
produced here:


```{r, warning=FALSE}

# Create survfit object
km.est.PFS <- survfit(Surv(PFS_days, PFS_event) ~ ARMCD, 
                      data = PFS_data, 
                      conf.type = 'plain')

# Plot Kaplan-Meier
plot(km.est.PFS, 
     col = c(blue, pink), # plot colours
     lty = c(1:2), # line type
     xlab = "Time (Days)", # x-axis label
     ylab = "Progression-free survival", # y-axis label
     xlim = c(0, 800)) 

legend(x = 500, 
       y = .9, 
       legend = c("Arm A", "Arm B"), 
       lty = c(1:2), 
       col = c(blue, pink))

```

# Fitting the models

The <tt>runPSM</tt> function fits parametric survival models for multiple
distributions using the <tt>flexsurv</tt> package (using the
<tt>flexsurvPlus</tt> function <tt>fit_models</tt>),  manipulates the flexsurv
objects to get the parameter estimates and AIC and BIC value (using the
<tt>flexsurvPlus</tt> function <tt>get_params</tt>) and rearranges the parameter
estimates such that they can easily be output to excel to calculate survival for
both the intervention and reference treatment in an economic model.

These functions can be used to estimate 3 types of model:
  
*  Separate models - Models fitted to each treatment arm separately
    *  These models are fitted using the <tt>runPSM</tt> function using the
    <tt>model.type="Separate"</tt> argument.

*  Independent shape models - Models fitted to both/all treatment arms including
a treatment covariate to model the effect of treatment on both the scale and
shape parameter(s) of the distribution.
    * They have been fit using the <tt>runPSM</tt> function using the 
    <tt>model.type="Independent shape"</tt> argument.

*  Common shape models - Models fitted to both/all treatment arms including a
treatment covariate to model the effect of treatment on the scale parameter of
the distribution. The shape parameter(s) of the distribution is common across
treatments which reflects an assumption of proportional hazards or accelerated
failure time between treatments depending on the distribution
    * These models have been fit using the <tt>runPSM</tt> function using  the
    <tt>model.type="Common shape"</tt> argument.
    
*  One arm models - Models fitted to the entire dataset (no treatment strata)
    *  These models are fitted using the <tt>runPSM</tt> function using the
    <tt>model.type="One arm"</tt> argument.


The inputs to the <tt>runPSM</tt> function are:

*  <tt>data</tt> - A data frame containing individual patient data for the
relevant time to event outcomes

*  <tt>time_var</tt> - Name of time variable in 'data'. Variable must be numerical
and >0.

*  <tt>event_var</tt> - Name of event variable in 'data'. Variable must be
numerical and contain 1's to indicate an event and 0 to indicate a censor.

*  <tt>model.type</tt> - Character vector indicating the types of model formula
provided. Permitted values are 'Common shape', 'Independent shape' or 'Separate'
as per the models explained above.

* <tt>distr</tt> - Each type of model can be fitted with multiple distributions.
The distributions available for this package are:

    * Exponential ('exp')

    * Weibull ('weibull')

    * Gompertz ('gompertz')

    * Log-normal ('lnorm')

    * Log-logistic ('llogis')

    * Generalized gamma ('gengamma')

    * Gamma ('gamma')

    * Generalized F ('genf')

*  <tt>strata_var</tt> - Name of stratification variable in "data". This is
usually the treatment variable and must be categorical. Not required when model.type='One arm'.

*  <tt>int_name</tt> - Character to indicate the name of the treatment of
interest, must be a level of the "strata_var" column in "data", used for
labelling the parameters.

*  <tt>ref_name </tt> - Character to indicate the name of the reference treatment,
must be a level of the "strata_var" column in "data", used for labelling the
parameters. Not required when model.type='One arm'.

More information about each function can be used by running the code <tt>?runPSM</tt>.

Example code to fit all two arm models is presented below.


```{r}
psm_PFS_all <- runPSM(data=PFS_data,
                     time_var="PFS_days",
                     event_var="PFS_event",
                     model.type= c("Common shape", 
                                   "Independent shape", 
                                   "Separate"),
                     distr = c('exp',
                               'weibull',
                               'gompertz',
                               'lnorm',
                               'llogis',
                               'gengamma',
                               'gamma',
                               'genf'),
                     strata_var = "ARMCD",
                     int_name="A",
                     ref_name = "B")
psm_PFS_all

```




# Estimate survival from the models and plot the curves

Survival at a given time, t, is estimated as follows:

$$ S(t) = P({T>t}) = 1 - F(t) $$

Where F(t) is the cumulative distribution function. 

To cross check survival estimates in Excel models, the following functions in R
can be used to estimate the cumulative distribution function at given time
points for each distribution explored in this package (the estimates from the
cumulative distribution function can then be subtracted from 1 to estimate the
survival probability):

* Exponential: <tt>pexp()</tt>

* Weibull: <tt>pweibull()</tt>

* Gompertz: <tt>pgompertz()</tt>

* Log-normal: <tt>plnorm()</tt>

* Log-logistic: <tt>pllogis()</tt>

* Generalized gamma: <tt>pgengamma()</tt>

* Gamma: <tt>pgamma()</tt>

* Generalized F: <tt>pgenf()</tt>

The parameters outputted from each of the fitted models are used as inputs to
these functions. The code below gives some examples.

```{r}
# Landmark survival

# vector of times to estimate survival (days)
landmark_times <- c(0, 100, 200, 300)


# Example 1: intervention arm, Weibull distribution, common shape model
surv_comshp_weibull_int <- 1 - pweibull(landmark_times,
                                        shape = psm_PFS_all$parameters_vector["comshp.weibull.shape.int"],
                                        scale = psm_PFS_all$parameters_vector["comshp.weibull.scale.int"])
surv_comshp_weibull_int


# Example 2: intervention arm, log-normal distribution, separate model
surv_sep_lnorm_int <- 1 - plnorm(landmark_times,
                                 meanlog = psm_PFS_all$parameters_vector["sep.lnorm.meanlog.int"],
                                 sdlog = psm_PFS_all$parameters_vector["sep.lnorm.sdlog.int"])
surv_sep_lnorm_int


# Example 3: reference arm, Generalized Gamma distribution, independent shape model
surv_indshp_gengamma_ref <- 1 - pgengamma(landmark_times,
                                          mu = psm_PFS_all$parameters_vector["indshp.gengamma.mu.ref"],
                                          sigma = psm_PFS_all$parameters_vector["indshp.gengamma.sigma.ref"],
                                          Q = psm_PFS_all$parameters_vector["indshp.gengamma.Q.ref"])
surv_indshp_gengamma_ref

```



The same functions can be used to generate the data required to plot the
survival curves, overlaid on top of the KM plot. The time argument should
reflect how long you want to the extrapolate for and the unit of time is the
same as the input data (in this example, days).


```{r}

# Plot the common shape models (Weibull distribution) with the Kaplan-Meier

# vector of times to estimate survival (days)
times <- c(seq(from = 0, to = 1000, by = 0.1))


# Survival probabilities: intervention arm, Weibull distribution, common shape model
surv_comshp_weibull_int <- 1 - pweibull(times,
                                  shape = psm_PFS_all$parameters_vector["comshp.weibull.shape.int"],
                                  scale = psm_PFS_all$parameters_vector["comshp.weibull.scale.int"])

# Survival probabilities: reference arm, Weibull distribution, common shape model
surv_comshp_weibull_ref <- 1 - pweibull(times,
                                  shape = psm_PFS_all$parameters_vector["comshp.weibull.shape.ref"],
                                  scale = psm_PFS_all$parameters_vector["comshp.weibull.scale.ref"])

# Create two data frames that include the survival probablaities and times
surv_comshp_weibull_int_times <- data.frame(Time = times,
                                            Surv = surv_comshp_weibull_int,
                                            Trt = "Intervention")
surv_comshp_weibull_ref_times <- data.frame(Time = times,
                                            Surv = surv_comshp_weibull_ref,
                                            Trt = "Reference")

# Plot Kaplan-Meier
plot(km.est.PFS, 
     col = c(blue, pink), # plot colours
     lty = c(1:2), # line type
     xlab = "Time (Days)", # x-axis label
     ylab = "Progression-free survival", # y-axis label
     xlim = c(0, 1000)) 

# Add legend
legend(x = 500, 
       y = .9, 
       legend = c("Arm A", "Arm B"), 
       lty = c(1:2), 
       col = c(blue, pink))

# Add model estimates
lines(x = surv_comshp_weibull_int_times$Time, y = surv_comshp_weibull_int_times$Surv, col = blue)
lines(x = surv_comshp_weibull_ref_times$Time, y = surv_comshp_weibull_ref_times$Surv, col = pink)

```


# Adressing uncertainty - Bootstrapping

Bootstrapping has been used to estimate the uncertainty of the parameters from
the survival models. Boostrapping is used for two reasons motivated by intent of
this package to support further modeling in excel.
1. To simplify and accelerate calculations in excel while maintaining
correlations between parameters (as is commonly done for NMA)
2. To maintain correlations across multiple endpoints (see separate vignette for
details)

Bootstrapping involves:

1. Sampling, with replacement, from all patients
2. Estimating all specified parametric survival models

This procedure is repeated multiple times to obtain a distribution of parameters.
For this example, bootstrap estimates of the parameters were calculated using the
<tt>boot</tt> package. An argument for the <tt>boot</tt> function is
<tt>statistic</tt> which is a function which when applied to data returns a
vector containing the statistic(s) of interest. The <tt>bootPSM</tt> function 
in the <tt>flexsurvPlus</tt> package can be used for this purpose.

The inputs for the <tt>bootPSM</tt> function are identical to the
<tt>runPSM</tt> function, however there is one additional argument:

*  <tt>i</tt> - Index used to select a sample within <tt>boot</tt>


```{r}
set.seed(2358)
# For speed and illustration only 4 models are selected for bootstrapping - all could be included. 
# To minimize vignette computation time only 100 bootstrap samples are taken. In general more samples should be used.
n.sim <- 100 

PSM_bootstraps_PFS <- boot(
  statistic = bootPSM, # bootstrap function
  R=n.sim, # number of bootstrap samples
  data=PFS_data,
  time_var="PFS_days",
  event_var="PFS_event",
  model.type = c("Common shape", "Separate"),
  distr = c('weibull', 'gamma'),
  strata_var = "ARMCD",
  int_name = "B",
  ref_name = "A"
)

```

To use the result of these samples it is helpful to do some post processing to make the resulting samples easier to interpret.


```{r}
# first extract the bootstrapped parameters into a tibble
PFS_bootsamples <- as_tibble(PSM_bootstraps_PFS$t)
# then add column names so can identify model and parameter more easily
colnames(PFS_bootsamples) <- names(PSM_bootstraps_PFS$t0)

# show the first 3 samples
PFS_bootsamples[1:3,]

```

# Estimating quantities from the sample

As the <tt>flexsurv</tt> parameterisations are used any quantity of interest can
be simply calculated for all models and samples through use of the flexsurv
functions such as the extrapolated means via flexsurv::mean_weibull. For this example we assume 
we have decided that separate models for reference and intervention are most appropriate and that
for the reference arm a gamma model is preferred while for the intervention arm a weibull model is best.

```{r}
# we can now calculate the mean PFS for the selected models for each bootstrap sample
# using weibull for the reference
PFS_ref_mean <- with(PFS_bootsamples, flexsurv::mean_gamma(shape = sep.gamma.shape.ref, rate = sep.gamma.rate.ref))
# using gamma for the intervention
PFS_int_mean <- with(PFS_bootsamples, flexsurv::mean_weibull(shape = sep.weibull.shape.int, scale = sep.weibull.scale.int))

# we can then calculate the incremental mean PFS from these
PFS_delta = PFS_int_mean - PFS_ref_mean 

# if we are interested we can then estimate quantiles from this
quantile(PFS_delta, probs = c(0.025,0.975))

# or plot the density of this derived quantity

density <- density(PFS_delta)
plot(density, 
     lwd = 2, 
     main = "Density")


```



# Outputing parameters to excel 

The primary use of the bootstrap samples is to be used in probabilistic
sensitivity analyses in economic models.

Once all the models have been fit and bootstrap samples estimated they can be
output to excel. By selecting the "Main Estimates" the estimates for the original data
are returned. To run PSA in excel only a random number between 1 and the number
of samples needs to be generated and the associated Bootstrap sample selected.

```{r}

# combine the estimates from the main analysis with the bootstrap samples
# and add meta data to include details of analysis

# first we can get combine the main estimates for the models with those that were bootstrapped
parameters_PFS <- rbind(PSM_bootstraps_PFS$t0, as_tibble(PSM_bootstraps_PFS$t))

# we can now add names
colnames(parameters_PFS) <- names(PSM_bootstraps_PFS$t0)

# we can then label the samples and add some metadata
metadata_PFS <- tibble(Estimate = c("Main Estimates", paste("Bootstrap Sample",1:n.sim)),
                       Study_name = "Study ABC",
                        Datacut = "Final",
                        Population = "ITT",
                        Endpoint = "PFS",
                        RefArmName = PSM_bootstraps_PFS$call$ref_name,
                        IntArmName = PSM_bootstraps_PFS$call$int_name)

pfs_for_export <- cbind(metadata_PFS, parameters_PFS)

# not run
#write.csv(parameters_PFS, "params_for_model.csv")

```


# References
