---
title: 'Example for a Markov Model'
author: "Javier Sanchez Alvarez and Valerie Aponte Ribero"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{example_markov}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
```


# Introduction
This document runs a discrete event simulation model in the context of a simple cohort Markov model with 4 states. Note that this same exercise could be done from a patient simulation approach rather than the cohort one.

## Main options

```{r setup}
library(descem)

library(dplyr)
library(flexsurv)
library(ggplot2)
library(kableExtra)
library(purrr)
library(tidyr)

#Show all numbers, no scientific notation
options(scipen = 999)


```

## Model Concept
The model is a simple Markov model with 4 states whose transition matrix can be found below. In order to run a pure Markov model within these functions, we will define each event as each cycle. We will generate an initial trace and at each event (cycle) we will update the trace by multiplying it by the transition matrix. Costs and QALYs can be computed in a similar fashion by multiplying the trace times the cost and the utility.

## Load Data
The dummy data is generated below. The data structure should be as defined below, otherwise it will give problems.

```{r load_data}
#Utilities
util.data <- data.frame( name = c("util1" ,"util2" ,"util3" ,"util4"),
                         value = c(0.9,0.75,0.6,0),
                         se=rep(0.02,4),
                         stringsAsFactors = FALSE
)


#Costs
cost.data <- data.frame( name = c("cost1" ,"cost2" ,"cost3" ,"cost4","cost_int"),
                         value = c(1000,3000,6000,0,1000),
                         stringsAsFactors = FALSE
) %>%
  mutate(se= value/5)

```

# General inputs with delayed execution
Initial inputs and flags that will be used in the model can be defined below. We can define inputs that are common to all patients (`common_all_inputs`) within a simulation, inputs that are unique to a patient independently of the treatment (e.g. natural death, defined in `common_pt_inputs`), and inputs that are unique to that patient and that treatment (`unique_pt_inputs`). Items can be included through the `add_item` function, and can be used in subsequent items. All these inputs are generated before the events and the reaction to events are executed. Furthermore, the program first executes `common_all_inputs`, then `common_pt_inputs` and then `unique_pt_inputs`. So one could use the items generated in `common_all_inputs` in `unique_pt_inputs`.

We also define here the specific utilities and costs that will be used in the model. It is strongly recommended to assign unnamed objects if they are going to be processed in the model. In this case, we're only using `util_v` and `cost_v` as an intermediate input and these objects will not be processed (we just use them to make the code more readable), so it's fine if we name them.

We define here our initial trace, the number of cycles to be simulated, the transition matrices and the initial cycle time (i.e. 0). 

```{r input_delayed}
#Put objects here that do not change on any patient or intervention loop, for example a HR
common_all_inputs <- add_item(max_n_cycles = 30) %>%
  add_item( #utilities
    util_v = if(psa_bool){
      setNames(draw_gamma(util.data$value,util.data$se^2),util.data$name) #in this case I choose a gamma distribution
    } else{setNames(util.data$value,util.data$name)},
    util1 = util_v[["util1"]], 
    util2 = util_v[["util2"]],
    util3 = util_v[["util3"]],
    util4 = util_v[["util4"]]) %>%
  add_item( #costs
    cost_v = if(psa_bool){
      setNames(draw_gamma(cost.data$value,cost.data$se),cost.data$name) #in this case I choose a gamma distribution
    } else{setNames(cost.data$value,cost.data$name)},
    cost1    = cost_v[["cost1"]],
    cost2    = cost_v[["cost2"]],
    cost3    = cost_v[["cost3"]],
    cost4    = cost_v[["cost4"]],
    cost_int = cost_v[["cost_int"]])


#Put objects here that change as we loop through treatments for each patient (e.g. events can affect fl.tx, but events do not affect nat.os.s)
#common across trt but changes per pt could be implemented here (if (trt==)... )
unique_pt_inputs <- add_item(
                            trace = c(1,0,0,0),
                            transition = if( trt=="noint"){ 
                                            matrix(c(0.4,0.3,0.2,0.1,
                                            0.1,0.4,0.3,0.2,
                                            0.1,0.1,0.5,0.3,
                                            0,0,0,1),nrow=4,byrow=T)
                                         } else{
                                            matrix(c(0.5,0.3,0.1,0.1,
                                                     0.2,0.4,0.3,0.1,
                                                     0.1,0.2,0.5,0.2,
                                                     0,0,0,1),nrow=4,byrow=T)
                                              },
                            cycle_time = 0
)

```

# Events
## Add Initial Events
In our model, the events are `start and cycle`.

```{r model_evts}
init_event_list <- 
  add_tte(trt="noint",evts=c("start","cycle"),input={ #intervention
    start <- 0
    
    cycle <- 1
    
  }) %>%
  add_tte(trt="int",evts=c("start","cycle"),input={ #intervention
    start <- 0
    
    cycle <- 1
   
    })

```

## Add Reaction to Those Events
The explanation on how these part works can be seen in the early breast cancer tutorial.

In this Markov model case, in the event `start` we generate as many cycles as we need. At each `cycle` event we update the time of the cycle to keep track of it when we produce the output of the model and we update the trace. Finally, when all the events are over, we finish the simulation by setting `curtime` to infinity.
```{r model_reaction}

evt_react_list <-
  add_reactevt(name_evt = "start",
               input = {
                 for (i in 2:max_n_cycles) {
                   new_event(list("cycle" = curtime + i))
                 }
                 
               }) %>%
  add_reactevt(name_evt = "cycle",
               input = {
                 modify_item(list("cycle_time" = cycle_time + 1)) #Update cycle time
                 
                 modify_item(list( "trace" = trace %*% transition)) #Update trace
                 
                 if (max_n_cycles == cycle_time) {
                   modify_item(list("curtime" = Inf)) #Indicate end of simulation for patient
                 }
               }) 


```

# Costs and Utilities
Costs and utilities are introduced below. However, it's worth noting that the model is able to run without costs or utilities. One would just need to define the all the utility and costs related objects as `NULL` and the model would automatically assume they take value `0`.

### Utilities
Utilities are defined using pipes with the `add_util` function. In this case case, we are just multiplying the trace times the utilities at each state.

```{r utilities}


util_ongoing <- add_util(evt = c("start","cycle"),
                              trt = c("int", "noint"),
                              util = trace %*% c(util1,util2,util3,util4)
) #common utility across arms

```
### Costs
Costs are defined using pipes with the `add_cost` function, in a similar fashion to the utilities. In this case we add the costs in a continuous fashion.

```{r costs}

cost_ongoing <-
  add_cost(
    evt = c("start","cycle") ,
    trt = "noint",
    cost = trace %*% c(cost1,cost2,cost3,cost4)) %>%
  add_cost(
    evt = c("start","cycle") ,
    trt = "int",
    cost = trace %*% c(cost1+ cost_int,cost2+ cost_int,cost3+ cost_int,cost4) ) 


```


# Model
## Model Execution
The model can be run using the function `RunSim` below. We must define the number of patients to be simulated, the number of simulations, whether we want to run a PSA or not, the strategy list, the inputs, events and reactions defined above, the number of cores to be used (by default uses 1 core), the discount rate for costs and the discount rate for qalys. It is recommended not to use all the cores in the machine.

It is worth noting that the `psa_bool` argument does not run a PSA automatically, but is rather an additional input/flag of the model that we use as a reference to determine whether we want to use a deterministic or stochastic input. As such, it could also be defined in `common_all_inputs` as the first item to be defined, and the result would be the same. However, we recommend it to be defined in `RunSim`.

Note that the distribution chosen, the number of events and the interaction between events can have a substantial impact on the running time of the model. Since we are taking a cohort approach, we just need to indicate `npats = 1`.

```{r model_run}
#Logic is: per patient, per intervention, per event, react to that event.
results <- RunSim(  
  npats=1,                               # number of patients, recommended to set to 1000 if using PSA as it takes quite a while
  n_sim=1,                                  # if >1, then PSA, otherwise deterministic
  psa_bool = FALSE,
  trt_list = c("int", "noint"),             # intervention list
  common_all_inputs = common_all_inputs,    # inputs common that do not change within a simulation
  unique_pt_inputs = unique_pt_inputs,      # inputs that change within a simulation between interventions
  init_event_list = init_event_list,        # initial event list
  evt_react_list = evt_react_list,          # reaction of events
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  ncores = 1,                               # number of cores to use, recommended not to use all
  drc = 0.035,                              # discount rate for costs
  drq = 0.035,                              # discount rate for QALYs
  input_out = c(                            # list of additional outputs (Flags, etc) that the user wants to export for each patient and event
    "trace",
    "cycle_time"
  )
)
```
# Post-processing of Model Outputs
## Summary of Results

Once the model has been run, we can use the results and summarize them using the `summary_results_det` to print the results of the last simulation (if nsim=1, it's the deterministic case), and `summary_results_psa` to show the PSA results (with the confidence intervals). We can also use the individual patient data generated by the simulation, which we collect here in the `psa_ipd` object.  Note that the data for life years is wrong, as the model assumes we are running a patient simulation data and therefore it's adding the 4 states, inflating the total life years. We can manually adjust this to get the correct life years.

```{r post-processing_summary}

summary_results_det(results$final_output) #will print the last simulation!

psa_ipd_simple <- bind_rows(map(results$output_psa, "merged_df")) %>% select(-trace) %>% distinct() 

psa_ipd <- bind_rows(map(results$output_psa, "merged_df")) 

trace_t <- psa_ipd %>% mutate(state = rep(seq(1:4),62)) %>% select(trt,trace,state,evttime) 

trace_t[1:10,] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

life_years <-  psa_ipd %>%
  mutate(state = rep(seq(1:4),62)) %>% 
  group_by(trt) %>%
  filter(state!=4) %>% #erase death state for LY computation
  mutate(ly_final = ly*lag(trace,3L)) %>% #multiply by previous trace
summarise(ly_final = sum(ly_final,na.rm = TRUE)) #get final discounted life years

life_years %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

results$final_output[["lys.int"]] <- life_years$ly_final[life_years$trt=="int"]
results$final_output[["lys.noint"]] <- life_years$ly_final[life_years$trt=="noint"]

summary_results_det(results$final_output) #will print the last simulation!

```

We can also check each of the cycles

```{r post-processing_analysis,echo=FALSE, message=FALSE}

psa_ipd_simple %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## Plots

We now use the data to plot the traces. 

```{r post-processing_plots1, fig.width=10, fig.height=8}

ggplot(trace_t,aes(x=evttime,y = trace,col=trt)) + geom_line() + facet_wrap(~state)

```
